---
title: "HPV Baseline Preliminary Data Analysis"
author: Myo Minn Oo & Joel Fowke
date: 2023-07-28
date-modified: last-modified
title-block-style: plain
title-block-banner: true
code-fold: true
code-copy: true
code-overflow: wrap
code-tools: true
toc: true
toc-depth: 3
toc-expand: true
format: html
---

```{r setup}
#| warning: false
#| message: false
here::here("data", "hpv_processed.rda") |> 
	load()

## load packages
library(tidyverse)
library(gtsummary)

## import custom-functions from utils
source(here::here("R", "utils.R"), echo = FALSE)

## relabelling
vars$sex <- "Cohort Type"

combined <- combined |> 
	mutate(
		age_cat = case_when(
			Age <= 20 ~ "≤20", 
			Age > 20 & Age <= 25 ~ "21-25", 
			Age > 25 & Age <= 30 ~ "25-30", 
			Age > 30 & Age <= 35 ~ "31-35", 
			Age > 35 & Age <= 40 ~ "36-40", 
			Age > 40 & Age <= 45 ~ "41-45", 
			Age > 45 & Age <= 50 ~ "46-50", 
			Age > 50 ~ "51+"
		), 
		age_cat = factor(age_cat), 
		across(hr_hpv_16:hr_p5, ~ 1/.x), 
		across(hr_hpv_16:hr_p5, ~ if_else(is.infinite(.x), NA, .x)), 
		sex_work_yr = case_when(
			sex_work_yr >= 0 & sex_work_yr <= 3 ~ "0-3", 
			sex_work_yr > 3 & sex_work_yr <= 6 ~ "3-6", 
			sex_work_yr > 6 ~ "7+", 
		)
	) |> 
	labelled::set_variable_labels(.labels = vars, .strict = FALSE)
```

## About

This is a preliminary data analysis using baseline data from Gates-funded HPV study in Nairobi, 2022-2023. The screening period is between May/June 2022 and July 2023.

## Research questions

1.  What is the prevalence of HPV among female sex workers and the distributions of genotypes (using geneXpert machine; hence five channel)?
2.  What baseline characteristics correlate with HPV prevalence?

## Known issues

The study screened 1,968 participants and enrolled 701. However, due to filing issues which the team in Nairobi are currently fixing, the database comprises of only `r nrow(combined)` unique individuals (short of `r 1968 - nrow(combined)`), which are used in the following analysis.

## Exploratory data analysis

### Comparing baseline characteristics between FSW and MSM:

```{r}
#| error: false
#| warning: false
#| message: false
tbl_merge(
	tbls = list(
		combined |> 
			dplyr::select(
				sex, Gender, Age, age_cat, hiv_self_report, 
				MaritalStatus, education_level
			) |> 
			table1_total(missing = "ifany"), 
		combined |> 
			dplyr::select(
				Gender, sex, Age, age_cat, hiv_self_report, 
				MaritalStatus, education_level
			) |> 
			table1_by(by = sex, 
						# missing = "ifany", 
						percent = "row", 
						add_N = FALSE, 
						add_total = FALSE)
	),
	tab_spanner = c(NA, "**Cohort Type**")
)
```

### Population Pyramid Graph

```{r}
#| warning: false
#| fig-width: 8
#| fig-height: 6
combined |> 
	apyramid::age_pyramid(age_cat, sex) + 
	labs(
		y = "Frequency", 
		x = "Age"
	) + 
	theme_bw()
```

## Prevalence of HPV by baseline characteristics

```{r}
#| error: false
#| warning: false
#| message: false
fsw <- 
	combined |> 
	dplyr::filter(sex == "FSW") |> 
	dplyr::mutate(
		Gender = forcats::fct_drop(Gender)
	)


gtsummary::tbl_merge(
	tbls = list(
		fsw |> 
			dplyr::select(sex, Gender:douching_with) |> 
			table1_total(missing = "ifany"), 
		fsw |> 
			dplyr::select(Gender:douching_with, hpv_result) |> 
			table1_by(by = hpv_result, byname = "HPV Result", 
						# missing = "ifany", 
						percent = "row", 
						add_N = FALSE, 
						add_total = FALSE)
	), 
	tab_spanner = c(NA, "HPV Result")
)
```

## Five genotype channels showing ct values

> P1 = HPV 16
>
> P2 = HPV 18 or 45
>
> P3 = HPV 31,33,35,52 or 58
>
> P4 = HPV 51 or 59
>
> P5 = HPV 39, 56, 66 or 68

```{r}
#| error: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5


d <- fsw |> 
	dplyr::filter(hpv_result == "Positive") |> 
	dplyr::select(hr_hpv_16:hr_p5) |> 
	setNames(paste0("P", 1:5)) |> 
	tidyr::pivot_longer(cols = everything()) |> 
	dplyr::filter(!is.na(value)) |> 
	dplyr::mutate(value = if_else(value == 0, NA, value))

d |> 
	ggplot(aes(x = name, y = value)) + 
	geom_boxplot(
		aes(
			fill = name, 
			fill = after_scale(colorspace::lighten(fill, .5)),
			color = after_scale(colorspace::darken(fill, .5)),
		), 
		width = .5, 
		show.legend = FALSE
	) + 
	geom_point(
		aes(
			fill = name, 
			fill = after_scale(colorspace::lighten(fill, .5)),
			color = after_scale(colorspace::darken(fill, .5)),
		), 
		position = position_jitter(.15, seed = 123), 
		shape = 21, 
		size = 2.5, 
		stroke = .5,
		alpha = .6
	) + 
	geom_text(
		data = 
			d |> 
			dplyr::filter(!is.na(value)) |> 
			dplyr::count(name) |> 
			dplyr::mutate(
				pct = n / sum(n), 
				n = paste0("(n=", n, ", ", scales::label_percent(accuracy = .1)(pct), ")"),
				value = 0.06
			), 
		aes(label = n)
	) + 
	labs(
		x = NULL, 
		y = "1/Ct value", 
		title = "Genotype channels showing 1/ct values among HPV positive patients"
	) + 
	theme_bw() + 
	theme(
		legend.position = "none", 
		plot.title = element_text(face = "bold"), 
		axis.title = element_text(face = "bold")
	)

```

```{r}
#| error: false
#| warning: false
#| message: false
#| fig-width: 16
#| fig-height: 12
d <- fsw |> 
	dplyr::filter(hpv_result == "Positive") |> 
	dplyr::select(hr_hpv_16:hr_p5, hiv_self_report, age_cat) |> 
	setNames(c(paste0("P", 1:5), "HIV", "Age")) |> 
	tidyr::pivot_longer(cols = -c(HIV, Age), names_to = "Genotype") |> 
	dplyr::filter(!is.na(value), !is.na(HIV)) |> 
	dplyr::mutate(value = if_else(value == 0, NA, value))

d |> 
	ggplot(aes(x = Age, y = value)) + 
	geom_boxplot(
		aes(
			fill = Age, 
			fill = after_scale(colorspace::lighten(fill, .5)),
			color = after_scale(colorspace::darken(fill, .5)),
		), 
		width = .5, 
		show.legend = FALSE
	) + 
	geom_point(
		aes(
			fill = Age, 
			fill = after_scale(colorspace::lighten(fill, .5)),
			color = after_scale(colorspace::darken(fill, .5)),
		), 
		position = position_jitter(.15, seed = 123), 
		shape = 21, 
		size = 2.5, 
		stroke = .5,
		alpha = .6
	) + 
	geom_text(
		data = 
			d |> 
			dplyr::filter(!is.na(value)) |> 
			dplyr::count(HIV, Age, Genotype) |> 
			dplyr::group_by(HIV, Age) |> 
			dplyr::mutate(
				pct = n / sum(n), 
				n = paste0("(n=", n, ", ", scales::label_percent(accuracy = .1)(pct), ")"),
				value = 0.06
			), 
		aes(label = n)
	) + 
	labs(
		x = NULL, 
		y = "Ct value", 
		title = "Genotype channels showing ct values among HPV positive patients"
	) + 
	theme_bw() + 
	theme(
		legend.position = "none", 
		plot.title = element_text(face = "bold"), 
		axis.title = element_text(face = "bold")
	) + 
	facet_grid(Genotype ~ HIV, labeller = label_both)

```

## Genotype distributions

```{r}
#| error: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5
fsw <- 
	fsw |> 
	dplyr::mutate(
		across(hr_hpv_16:hr_p5, 
					 ~ if_else(
					 	hpv_result == "Positive" & !is.na(.x), "Yes", "No" 
					 )), 
		hpv_combined = (hr_hpv_16 == "Yes") + (hr_hpv_18_45 == "Yes") + 
			(hr_p3 == "Yes") + (hr_p4 == "Yes") + (hr_p5 == "Yes"), 
		hpv_combined = if_else(hpv_combined == 0, NA, hpv_combined)
	) |> 
	labelled::set_variable_labels(.labels = vars, .strict = FALSE)


gtsummary::tbl_merge(
	tbls = list(
		fsw |>
			dplyr::select(hr_hpv_16:hpv_combined, hiv_self_report) |>
			table1_total(missing = "ifany"),
		fsw |>
			dplyr::select(hpv_result, hr_hpv_16:hpv_combined) |>
			table1_by(by = hpv_result, byname = "HPV Result",
						# missing = "ifany",
						percent = "row",
						add_N = FALSE,
						add_total = FALSE)
	),
	tab_spanner = c(NA, "HPV Result")
)


fsw |>
			dplyr::select(hpv_combined, hiv_self_report) |>
			table1_by(by = hiv_self_report, byname = "HIV Result",
						# missing = "ifany",
						percent = "row",
						add_N = FALSE,
						add_total = FALSE)
	
fsw |> 
	dplyr::select(hr_hpv_16:hr_p5) |> 
	setNames(paste0("P", 1:5)) |> 
	tidyr::pivot_longer(cols = everything()) |> 
	dplyr::filter(!is.na(value)) |> 
	dplyr::group_by(name, value) |> 
	dplyr::summarise(n = n()) |> 
	dplyr::mutate(
		pct = n / sum(n), 
		pct = paste0(n, " (", scales::label_percent()(pct), ")")
	) |> 
	dplyr::ungroup() |> 
	dplyr::filter(value != "No") |> 
	ggplot(aes(y = name, x = n)) + 
	geom_col(aes(fill = name), show.legend = FALSE) + 
	geom_label(aes(label = pct), hjust = 0) + 
	labs(
		x = "Number", 
		y = "Genotype Channel", 
		title = "HPV Genotype Distribution among FSW"
	) + 
	scale_x_continuous(
		expand = c(.01, .01), 
		limits = c(0, 400)
		
	) + 
	theme_bw() +
	theme(
		plot.title = element_text(face = "bold"), 
		axis.title = element_text(face = "bold")
	)

fsw |> 
	dplyr::select(hr_hpv_16:hr_p5, hiv_self_report) |> 
	setNames(c(paste0("P", 1:5), "HIV")) |> 
	tidyr::pivot_longer(cols = -HIV) |> 
	dplyr::filter(!is.na(value)) |> 
	dplyr::group_by(HIV, name, value) |> 
	dplyr::summarise(n = n()) |> 
	dplyr::mutate(
		pct = n / sum(n), 
		label = paste0(n, " (", scales::label_percent()(pct), ")")
	) |> 
	dplyr::ungroup() |> 
	dplyr::filter(value != "No") |> 
	dplyr::filter(!is.na(HIV)) |> 
	ggplot(aes(y = name, x = pct)) + 
	geom_col(aes(fill = name), show.legend = FALSE) + 
	geom_label(aes(label = label), hjust = 0) + 
	labs(
		x = "Number", 
		y = "Genotype Channel", 
		title = "HPV Genotype Distribution by HIV Status"
	) + 
	scale_x_continuous(
		expand = c(.01, .01), 
		labels = scales::label_percent(),
		limits = c(0, .5)
	) + 
	facet_wrap(~ HIV, labeller = label_both) +
	theme_bw() +
	theme(
		plot.title = element_text(face = "bold"), 
		axis.title = element_text(face = "bold")
	)


```

```{r}
#| error: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 16

fsw |> 
	dplyr::select(hr_hpv_16:hr_p5, hiv_self_report, age_cat) |> 
	setNames(c(paste0("P", 1:5), "HIV", "Age")) |> 
	tidyr::pivot_longer(cols = -c(HIV, Age)) |> 
	dplyr::filter(!is.na(value)) |> 
	dplyr::group_by(Age, HIV, name, value) |> 
	dplyr::summarise(n = n()) |> 
	dplyr::mutate(
		pct = n / sum(n), 
		label = paste0(n, " (", scales::label_percent()(pct), ")")
	) |> 
	dplyr::ungroup() |> 
	dplyr::filter(value != "No") |> 
	dplyr::filter(!is.na(Age), !is.na(HIV)) |> 
	ggplot(aes(y = name, x = pct)) + 
	geom_col(aes(fill = name), show.legend = FALSE) + 
	geom_label(aes(label = label), hjust = 0) + 
	labs(
		x = "Number", 
		y = "Genotype Channel", 
		title = "HPV Genotype Distribution by HIV Status"
	) + 
	scale_x_continuous(
		expand = c(.01, .01), 
		labels = scales::label_percent(),
		limits = c(0, .7)
	) + 
	facet_grid(Age ~ HIV, labeller = label_both) +
	theme_bw() +
	theme(
		plot.title = element_text(face = "bold"), 
		axis.title = element_text(face = "bold")
	)
```

## Number of HPV genotypes by HIV and HPV status

```{r}
#| error: false
#| warning: false
#| message: false
tbl_merge(
	tbls = list(
		fsw |> 
			dplyr::select(
				Gender, sex, Age, age_cat, hiv_self_report, 
				MaritalStatus, education_level
			) |> 
			table1_total(missing = "ifany"), 
		fsw |> 
			dplyr::select(
				Gender, sex, Age, age_cat, hiv_self_report, 
				MaritalStatus, education_level, hpv_hiv
			) |> 
			table1_by(by = hpv_hiv, 
						byname = "Combined HIV & HPV Status ", 
						# missing = "ifany", 
						percent = "row", 
						add_N = FALSE, 
						add_total = FALSE)
	),
	tab_spanner = c(NA, "**Combined HIV & HPV Status**")
)
	
```

## Factors associated with HPV

### Unadjusted Odds Ratios

```{r}
#| error: false
#| warning: false
#| message: false
fsw |> 
	dplyr::select(
		Gender, age_cat, hiv_self_report, 
		MaritalStatus:menopause, contraceptive, hpv_test,
		condom_lastsex:num_partner_reg, 
		num_partner_hiv_cas:num_partner_hiv_reg, 
		drug_sex_cas:drug_sex_reg, douching, douching_with,
		response
	) |> 
	dplyr::mutate(
		contraceptive = as.numeric(contraceptive != "None")
	) |>
	labelled::set_variable_labels(.labels = vars, .strict = FALSE) |> 
	tbl_uvregression(
		method = glm,
		y = response,
		method.args = list(family = binomial),
		exponentiate = TRUE
	)
```

### Adjusted Odds Ratios

```{r}
#| error: false
#| warning: false
#| message: false


mod_lm <- glm(
	response ~ Gender + age_cat + hiv_self_report +
		hpv_test + condom_access + sex_work_yr + num_partner_hiv_reg,
	family = binomial, 
	data = fsw
) 

mod_lm |> 
	gtsummary::tbl_regression(
		exponentiate = TRUE
	)
```

## TO-DO Lists

-   Explore more variables: sexual practices + egocentric variables
-   Other variables: violence, quality of life,
-   Integrate with flow data + immunological data

On technical side

-   Missing values
-   Could add more works on model selection and refining
