---
title: "HPV Baseline Preliminary Data Analysis"
author: Myo Minn Oo & Joel Fowke
date: last-modified
title-block-style: plain
title-block-banner: true
code-fold: true
code-copy: true
code-overflow: wrap
code-tools: true
toc: true
toc-depth: 3
toc-expand: false
format: html
---

```{r setup}
#| warning: false
#| message: false
here::here("data", "hpv_processed.rda") |> 
	load()

## load packages
library(tidyverse)
library(gtsummary)

## import custom-functions from utils
source(here::here("R", "utils.R"), echo = FALSE)
```

## About

This is a preliminary data analysis using baseline data from Gates-funded HPV study in Nairobi, 2022-2023. The screening period is between May/June 2022 and July 2023.

## Research questions

1.  What is the prevalence of HPV among female sex workers and the distributions of genotypes (using geneXpert machine; hence five channel)?
2.  What baseline characteristics correlate with HPV prevalence?

## Known issues

The study screened 1,968 participants and enrolled 701. However, due to filing issues which the team in Nairobi are currently fixing, the database comprises of only `r nrow(combined)` unique individuals (short of `r 1968 - nrow(combined)`), which are used in the following analysis. 

## Exploratory data analysis

### Comparing baseline characteristics between FSW and MSM:

```{r}
#| error: false
#| warning: false
#| message: false
tbl_merge(
	tbls = list(
		combined |> 
			dplyr::select(
				Gender, sex, Age, age_cat, hiv_self_report, 
				MaritalStatus, education_level
			) |> 
			table1_total(missing = "ifany"), 
		combined |> 
			dplyr::select(
				Gender, sex, Age, age_cat, hiv_self_report, 
				MaritalStatus, education_level
			) |> 
			table1_by(by = sex, 
						byname = "Participant Type by Sex", 
						# missing = "ifany", 
						percent = "row", 
						add_N = FALSE, 
						add_total = FALSE)
	),
	tab_spanner = c(NA, "**Participant Type by Sex**")
)

```


### Population Pyramid Graph 

```{r}
#| warning: false
#| fig-width: 8
#| fig-height: 8
combined |> 
	dplyr::mutate(
		age_cat2 = cut(Age, c(0, 20, 30, 40, 50, 60, 100))
	) |> 
	apyramid::age_pyramid(age_cat2, sex) + 
	labs(
		y = "Frequency", 
		x = "Age"
	) + 
	theme_classic()
```




## Prevalence of HPV by baseline characteristics

```{r}
#| error: false
#| warning: false
#| message: false
fsw <- 
	combined |> 
	dplyr::filter(sex == "FSW") |> 
	dplyr::mutate(
		Gender = forcats::fct_drop(Gender)
	)


gtsummary::tbl_merge(
	tbls = list(
		fsw |> 
			dplyr::select(Gender:douching_with) |> 
			table1_total(missing = "ifany"), 
		fsw |> 
			dplyr::select(Gender:douching_with, hpv_result) |> 
			table1_by(by = hpv_result, byname = "HPV Result", 
						# missing = "ifany", 
						percent = "row", 
						add_N = FALSE, 
						add_total = FALSE)
	), 
	tab_spanner = c(NA, "HPV Result")
)
```


## Genotype distributions

```{r}
#| error: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5
# gtsummary::tbl_merge(
# 	tbls = list(
# 		fsw |> 
# 			dplyr::select(hr_hpv_16:hpv_combined) |> 
# 			table1_total(missing = "ifany"), 
# 		fsw |> 
# 			dplyr::select(hr_hpv_16:hpv_combined, hpv_result) |> 
# 			table1_by(by = hpv_result, byname = "HPV Result", 
# 						# missing = "ifany", 
# 						percent = "row", 
# 						add_N = FALSE, 
# 						add_total = FALSE)
# 	), 
# 	tab_spanner = c(NA, "HPV Result")
# )
fsw |> 
	dplyr::select(hr_hpv_16:hr_p5) |> 
	setNames(paste0("P", 1:5)) |> 
	tidyr::pivot_longer(cols = everything()) |> 
	dplyr::filter(!is.na(value)) |> 
	dplyr::group_by(name, value) |> 
	dplyr::summarise(n = n()) |> 
	dplyr::mutate(
		pct = n / sum(n), 
		pct = paste0(n, " (", scales::label_percent()(pct), ")")
	) |> 
	dplyr::ungroup() |> 
	dplyr::filter(value != "No") |> 
	ggplot(aes(y = name, x = n)) + 
	geom_col(aes(fill = name), show.legend = FALSE) + 
	geom_label(aes(label = pct), hjust = 0) + 
	labs(
		x = "Number", 
		y = "Genotype Channel", 
		title = "HPV Genotype Distribution among FSW"
	) + 
	scale_x_continuous(
		expand = c(.01, .01), 
		
	) + 
	theme_classic() +
	theme(
		plot.title = element_text(face = "bold"), 
		axis.title = element_text(face = "bold")
	)

fsw |> 
	dplyr::select(hr_hpv_16:hr_p5, hiv_self_report) |> 
	setNames(c(paste0("P", 1:5), "HIV")) |> 
	tidyr::pivot_longer(cols = -HIV) |> 
	dplyr::filter(!is.na(value)) |> 
	dplyr::group_by(HIV, name, value) |> 
	dplyr::summarise(n = n()) |> 
	dplyr::mutate(
		pct = n / sum(n), 
		pct = paste0(n, " (", scales::label_percent()(pct), ")")
	) |> 
	dplyr::ungroup() |> 
	dplyr::filter(value != "No") |> 
	dplyr::filter(!is.na(HIV)) |> 
	ggplot(aes(y = name, x = n)) + 
	geom_col(aes(fill = name), show.legend = FALSE) + 
	geom_label(aes(label = pct), hjust = 0) + 
	labs(
		x = "Number", 
		y = "Genotype Channel", 
		title = "HPV Genotype Distribution by HIV Status"
	) + 
	scale_x_continuous(
		expand = c(.01, .01), 
		limits = c(0, 320)
	) + 
	facet_wrap(~ HIV, labeller = label_both) +
	theme_classic() +
	theme(
		plot.title = element_text(face = "bold"), 
		axis.title = element_text(face = "bold")
	)


fsw |> 
	dplyr::select(hr_hpv_16:hr_p5, hiv_self_report, age_cat) |> 
	setNames(c(paste0("P", 1:5), "HIV", "Age")) |> 
	tidyr::pivot_longer(cols = -c(HIV, Age)) |> 
	dplyr::filter(!is.na(value)) |> 
	dplyr::group_by(Age, HIV, name, value) |> 
	dplyr::summarise(n = n()) |> 
	dplyr::mutate(
		pct = n / sum(n), 
		pct = paste0(n, " (", scales::label_percent()(pct), ")")
	) |> 
	dplyr::ungroup() |> 
	dplyr::filter(value != "No") |> 
	dplyr::filter(!is.na(Age), !is.na(HIV)) |> 
	ggplot(aes(y = name, x = n)) + 
	geom_col(aes(fill = name), show.legend = FALSE) + 
	geom_label(aes(label = pct), hjust = 0) + 
	labs(
		x = "Number", 
		y = "Genotype Channel", 
		title = "HPV Genotype Distribution by HIV Status"
	) + 
	scale_x_continuous(
		expand = c(.01, .01), 
		limits = c(0, 220)
	) + 
	facet_grid(Age ~ HIV, labeller = label_both) +
	theme_classic() +
	theme(
		plot.title = element_text(face = "bold"), 
		axis.title = element_text(face = "bold")
	)
```

## Number of HPV genotypes by HIV and HPV status

```{r}
#| error: false
#| warning: false
#| message: false
tbl_merge(
	tbls = list(
		fsw |> 
			dplyr::select(
				Gender, sex, Age, age_cat, hiv_self_report, 
				MaritalStatus, education_level
			) |> 
			table1_total(missing = "ifany"), 
		fsw |> 
			dplyr::select(
				Gender, sex, Age, age_cat, hiv_self_report, 
				MaritalStatus, education_level, hpv_hiv
			) |> 
			table1_by(by = hpv_hiv, 
						byname = "Combined HIV & HPV Status ", 
						# missing = "ifany", 
						percent = "row", 
						add_N = FALSE, 
						add_total = FALSE)
	),
	tab_spanner = c(NA, "**Combined HIV & HPV Status**")
)
	
```



## Factors associated with HPV

### Unadjusted Odds Ratios
```{r}
#| error: false
#| warning: false
#| message: false
fsw |> 
	dplyr::select(
		Gender, age_cat, hiv_self_report, 
		MaritalStatus:menopause, contraceptive, hpv_test,
		condom_lastsex:num_partner_reg, 
		num_partner_hiv_cas:num_partner_hiv_reg, 
		drug_sex_cas:drug_sex_reg, douching, douching_with,
		response
	) |> 
	dplyr::mutate(
		contraceptive = as.numeric(contraceptive != "None")
	) |>
	labelled::set_variable_labels(.labels = vars, .strict = FALSE) |> 
	tbl_uvregression(
		method = glm,
		y = response,
		method.args = list(family = binomial),
		exponentiate = TRUE
	)
```



### Adjusted Odds Ratios

```{r}
#| error: false
#| warning: false
#| message: false


mod_lm <- glm(
	response ~ Gender + age_cat + hiv_self_report +
		hpv_test + condom_access + sex_work_yr + num_partner_hiv_reg,
	family = binomial, 
	data = fsw
) 

mod_lm |> 
	gtsummary::tbl_regression(
		exponentiate = TRUE
	)
```



## TO-DO Lists

- Explore more variables: sexual practices + egocentric variables
- Other variables: violence, quality of life, 
- Integrate with flow data + immunological data

On technical side

- Missing values
- Could add more works on model selection and refining